The vertical Fermi and eROSITA bubbles inclined jets: A proof-of-concept study

### 2021-12-27 paper.tex
update

  \subsection{The Galactic and Disk Models}
  As a proof-of-concept study, we approximate conventionally axisymmetric stellar potential of Milky Way\
  by an plane-parallel potential that is symmetric about the mid-plane $z=0$\
  in a simulation box size of\
  $14\times14\times28$ kpc, slightly larger than the size of eROISTA bubbles.

@@

  We tabulate parameters in the first four\
  categories of \Cref{table-parameters},\
  except for $\rho_{\text{atmp}}^{\text{peak}}$ that\
  can be derived from the known parameters and pressure equilibrium condition\
  on the interfaces $(z=\pm z_{0})$ between the disk and atmosphere.\

@@

  pyFC randomly generates dimensionless 3D scalar field $f(\bold{x})$\ % that is clumpy and porous.
  that obeys the log-normal probability distribution\
  with mean $\mu$ and dispersion $\sigma$,\
  and follows the power-law Kolmogorov spectrum
  \begin{equation}
    D(\bold{k})=\int k^{2} \hat{f}(\bold{k})\hat{f}^{*}(\bold{k})d\Omega \propto k^{-\beta},
    \label{Kolmogorov-spectrum}
  \end{equation}
  where $\hat{f}(\bold{k})$ is the Fourier transform of $f(\bold{x})$.\
  The spectrum $D(\bold{k})$ in the Fourier space is characterized by the power law index $\beta=5/3$,\
  the Nyquist limit $k_{\text{max}}$, and the lower cutoff wave number $k_{\text{min}}$.
  $k_{\text{max}}$ is one-half of the spatial resolution within disk,\
  and $k_{\text{min}}$ is 375.0, corresponding to the maximum size of an individual clump $\sim 20$ pc.\

  \citet{LA2002} and \citet{Wagner2012} have outlined a detailed procedure\
  for constructing a clumpy scalar field, and we do not repeat here.

@@

  We summarize the parameters of clumpy disk and their references in \Cref{table-parameters}.

  The AMR base level is covered by $16\times16\times32$ root cells, refined progressively on the mid-plane $z=0$\
  based on the gradient of mass desnity.\
  The highest refinement level is 7 so that the clumps within the disk can be adequately\
  resolved by approximately 30 cells along their diameter, 20 pc.

@@

\begin{enumerate}
  \item Assumptions on gravity:
    \begin{enumerate}
      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
            supported by gas pressure and by under its own self-gravity,\
            to mimic the realistic potential of Galactic bulge.
      \item The ISM disk and atmosphere are subjected to the fixed external potential\
            due to a disk bulge and dark matter halo.
      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
    \end{enumerate}
\end{enumerate}

\begin{enumerate}
\item Fixed external gravitational potential:
  \begin{enumerate}
    \item Bulge potential:
      \begin{enumerate}
       \item Peak density: $\rho_{\text{bulge}}^{\text{peak}}=4\times 10^{-24}$ g/cm$^3$.
       \item Potential (isothermal slab):
         \begin{equation}
           \Phi_{\text{bulge}}=\
           2\sigma^2_{\text{bulge}}\
           \ln\cosh\left(z\sqrt{\frac{2\pi G\rho_{\text{bulge}}^{\text{peak}}}{\sigma^2_{\text{bulge}}}}\right),
         \end{equation}
          where $\displaystyle \sigma_{\text{bulge}}=\
                \sqrt{\frac{k_{B}T_{\text{bulge}}}{m}}=100$ km/s \citep{velocity-dispersion-MW}.
      \end{enumerate}

    \item Dark logarithmic halo potential:
         \begin{equation}
           \Phi_{\text{halo}}=v^2_{\text{halo}}\ln\left(z^2+d^2_{\text{h}}\right),
         \end{equation}
         where $v_{\text{halo}}=131.5$ km/s, $d_{\text{h}}=12$ kpc \citep{Yang2013}.

    \item Total potential: $\Phi_{\text{total}}=\Phi_{\text{bulge}}+\Phi_{\text{halo}}$.
  \end{enumerate}

\item Clumpy cold disk:
  \begin{enumerate}
    \item Dimension: $14\times14\times0.2$ kpc. i.e. Scale height, $z_{0}=100$ pc. \citep{peak-ism-density}
    \item Peak mass density: $\rho_{\text{disk}}^{\text{peak}}=10^{-23}$ g/cm$^3$. \citep{peak-ism-density}
    \item Average temperature: $T_{\text{disk}}=10^{3}$ K. \citep{peak-ism-density}
    \item Average mass density:
          \begin{equation}
             \rho_{\text{disk}}=\rho_{\text{disk}}^{\text{peak}}
             \exp\left[-\frac{\Phi_{\text{total}}}{k_{B}T_{\text{disk}}/m}\right].
             \label{disk-density}
          \end{equation}

    \item Pressure: $\displaystyle p_{\text{disk}}=\rho_{\text{disk}}T_{\text{disk}}$.

    \item The clumpy cuboid, using the publicly\
          available pyFC code\footnote{\url{https://pypi.python.org/pypi/pyFC}},\
          are described by a log-normal distribution with mean 1.0 and variance 5.0.\
          Also, the power spectrum of the clumpy cuboid is characterized by $k_{\text{min}}$ and $\beta$,\
          where $k_{\text{min}}$ sets the maximum cloud size within the clumpy cuboid,\
          and $\beta$ is the slope of power spectrum in Fourier space.
          In this paper, we set $\beta=-5/3$ and $k_{\text{min}}=375$ to follow the Kolmogorov spectrum\
          and to limits the maximum size of an individual cloud to approximately 25 pc.

    \item The clumpy disk is then constructed by multiplying a fractal cuboid with \Cref{disk-density}.

    \item The clouds and voids within clumpy disk are pressure balanced.
  \end{enumerate}

\item Isothermal atmosphere:
  \begin{enumerate}
     \item The region outside the clumpy cold disk is atmosphere.
     \item $T_{\text{atmp}}=10^{6}$ K. \citep{temperature-MW}
     \item Density:
          \begin{equation}
             \rho_{\text{atmp}}=\rho_{\text{atmp}}^{\text{peak}}
             \exp\left[-\frac{\Phi_{\text{total}}}{k_{B}T_{\text{atmp}}/m}\right],
             \label{atmosphere-density}
          \end{equation}
          where $\rho_{\text{atmp}}^{\text{peak}}$ can be obtained by assuming that pressure\
          and gravitational potential are continuous at the interface ($z=\pm z_{0}$) between disk and atmosphere.
  \end{enumerate}

\end{enumerate}

\subsection{Cold disk}

>>>>

  \subsection{The Galactic and Disk Models}
  As a proof-of-concept study, we approximate conventionally axisymmetric stellar potential of Milky Way\
  by a plane-parallel potential that is symmetric about the mid-plane $z=0$\
  in a simulation box size of\
  $14\times14\times28$ kpc, slightly larger than the size of eROISTA bubbles.

@@

  We tabulate parameters in the first four\
  categories of \Cref{table-parameters},\
  except for $\rho_{\text{atmp}}^{\text{peak}}$ that\
  can be derived from the other known parameters and pressure equilibrium condition\
  on the interfaces $(z=\pm z_{0})$ between the disk and atmosphere.\

@@

  pyFC randomly generates dimensionless 3D scalar field $f(\bold{x})$\ % that is clumpy and porous.
  that obeys the log-normal probability distribution\
  with mean $\mu$ and dispersion $\sigma$,\
  and follows the power-law Kolmogorov spectrum
  \begin{equation}
    D(\bold{k})=\int k^{2} \hat{f}(\bold{k})\hat{f}^{*}(\bold{k})d\Omega \propto k^{-\beta},
    \label{Kolmogorov-spectrum}
  \end{equation}
  where $\hat{f}(\bold{k})$ is the Fourier transform of $f(\bold{x})$.\
  The spectrum $D(\bold{k})$ in the Fourier space is characterized by the power law index $\beta=5/3$,\
  the Nyquist limit $k_{\text{max}}$, and the lower cutoff wave number $k_{\text{min}}$.
  $k_{\text{max}}$ is one-half of the spatial resolution within disk,\
  and $k_{\text{min}}$ is 375.0, corresponding to the maximum size of an individual clump $\sim 20$ pc.\
  \citet{LA2002} and \citet{Wagner2012} have outlined a detailed procedure\
  for constructing a clumpy scalar field, and we do not repeat here.

@@

  \Cref{table-parameters} summarize the parameters of the clumpy disk and their references.

  On the basis of this setup, we cover the AMR base level with\
  $16\times16\times32$ root cells, refined progressively on the mid-plane $z=0$\
  based on the gradient of mass density.\
  We also restrict the refinement level is 7 within the cold disk so that\
  a molecular cloud can be adequately resolved by approximately 30 cells along their diameter, 20 pc.

@@

%\begin{enumerate}
%  \item Assumptions on gravity:
%    \begin{enumerate}
%      \item We use the potential of isothermal slab, symmetric about $z=0$ and\
%            supported by gas pressure and by under its own self-gravity,\
%            to mimic the realistic potential of Galactic bulge.
%      \item The ISM disk and atmosphere are subjected to the fixed external potential\
%            due to a disk bulge and dark matter halo.
%      \item In addition to the gravitational interaction, we ignore other interactions between stars and gases.
%      \item We also ignore the self-gravity of the ISM disk and of the atmosphere.
%      \item We ignore the centrifugal force of Milky Way rotation acting on the bubbles.
%      \item We use the potential of isothermal slab to mimic the gravitational potential due to the stellar bulge.
%      \item The interface between cold ISM disk and atmosphere is parallel to the Galactic plane and is pressure balanced.
%    \end{enumerate}
%\end{enumerate}
%
%
%
%\begin{enumerate}
%\item Fixed external gravitational potential:
%  \begin{enumerate}
%    \item Bulge potential:
%      \begin{enumerate}
%       \item Peak density: $\rho_{\text{bulge}}^{\text{peak}}=4\times 10^{-24}$ g/cm$^3$.
%       \item Potential (isothermal slab):
%         \begin{equation}
%           \Phi_{\text{bulge}}=\
%           2\sigma^2_{\text{bulge}}\
%           \ln\cosh\left(z\sqrt{\frac{2\pi G\rho_{\text{bulge}}^{\text{peak}}}{\sigma^2_{\text{bulge}}}}\right),
%         \end{equation}
%          where $\displaystyle \sigma_{\text{bulge}}=\
%                \sqrt{\frac{k_{B}T_{\text{bulge}}}{m}}=100$ km/s \citep{velocity-dispersion-MW}.
%      \end{enumerate}
%
%    \item Dark logarithmic halo potential:
%         \begin{equation}
%           \Phi_{\text{halo}}=v^2_{\text{halo}}\ln\left(z^2+d^2_{\text{h}}\right),
%         \end{equation}
%         where $v_{\text{halo}}=131.5$ km/s, $d_{\text{h}}=12$ kpc \citep{Yang2013}.
%
%    \item Total potential: $\Phi_{\text{total}}=\Phi_{\text{bulge}}+\Phi_{\text{halo}}$.
%  \end{enumerate}
%
%
%\item Clumpy cold disk:
%  \begin{enumerate}
%    \item Dimension: $14\times14\times0.2$ kpc. i.e. Scale height, $z_{0}=100$ pc. \citep{peak-ism-density}
%    \item Peak mass density: $\rho_{\text{disk}}^{\text{peak}}=10^{-23}$ g/cm$^3$. \citep{peak-ism-density}
%    \item Average temperature: $T_{\text{disk}}=10^{3}$ K. \citep{peak-ism-density}
%    \item Average mass density:
%          \begin{equation}
%             \rho_{\text{disk}}=\rho_{\text{disk}}^{\text{peak}}
%             \exp\left[-\frac{\Phi_{\text{total}}}{k_{B}T_{\text{disk}}/m}\right].
%             \label{disk-density}
%          \end{equation}
%
%    \item Pressure: $\displaystyle p_{\text{disk}}=\rho_{\text{disk}}T_{\text{disk}}$.
%
%    \item The clumpy cuboid, using the publicly\
%          available pyFC code\footnote{\url{https://pypi.python.org/pypi/pyFC}},\
%          are described by a log-normal distribution with mean 1.0 and variance 5.0.\
%          Also, the power spectrum of the clumpy cuboid is characterized by $k_{\text{min}}$ and $\beta$,\
%          where $k_{\text{min}}$ sets the maximum cloud size within the clumpy cuboid,\
%          and $\beta$ is the slope of power spectrum in Fourier space.
%          In this paper, we set $\beta=-5/3$ and $k_{\text{min}}=375$ to follow the Kolmogorov spectrum\
%          and to limits the maximum size of an individual cloud to approximately 25 pc.
%
%    \item The clumpy disk is then constructed by multiplying a fractal cuboid with \Cref{disk-density}.
%
%    \item The clouds and voids within clumpy disk are pressure balanced.
%  \end{enumerate}
%
%
%\item Isothermal atmosphere:
%  \begin{enumerate}
%     \item The region outside the clumpy cold disk is atmosphere.
%     \item $T_{\text{atmp}}=10^{6}$ K. \citep{temperature-MW}
%     \item Density:
%          \begin{equation}
%             \rho_{\text{atmp}}=\rho_{\text{atmp}}^{\text{peak}}
%             \exp\left[-\frac{\Phi_{\text{total}}}{k_{B}T_{\text{atmp}}/m}\right],
%             \label{atmosphere-density}
%          \end{equation}
%          where $\rho_{\text{atmp}}^{\text{peak}}$ can be obtained by assuming that pressure\
%          and gravitational potential are continuous at the interface ($z=\pm z_{0}$) between disk and atmosphere.
%  \end{enumerate}
%
%
%
%\end{enumerate}
%
